---
title: "MA [46]15 Homework 2"
author: "Prisma Lopez"
output: html_document
---

In this homework I'm performing a more detailed analysis of the `nycflights13` dataset I saw in lecture. I start by loading the packages:

```{r init}
library(tidyverse)
library(nycflights13)
```


## Question 1

Here are the top 4 carriers in total number of flights departing NYC in 2013:

**`[==[`** Change the `q1` chunk to create a dataset with only the top **4** carriers in total number of flights, then store these carriers in variable `big_carriers`. Recall that `dataset$variable` extracts `variable` values from `dataset`. **`]==]`**

```{r q1}
# original code:
#unique(flights$carrier) # all carriers
#(big_carriers <- c("AA", "UA", "DL"))
f <- count(flights,carrier)
f <- as.data.frame(f)

big_carriers_summary <- f %>% mutate(rank = min_rank(desc(n))) %>% arrange(rank) %>% filter(rank <=4) %>% select(carrier,rank,n)

big_carriers <- big_carriers_summary$carrier
big_carriers # outputs character string of all top four carriers in ascending rank order
```


## Question 2

Here are the proportion of canceled flights originated at each NYC airport for the top 4 carriers above:

```{r q2}
# SUBMIT:
flights %>% filter(carrier %in% big_carriers) %>% 
  group_by(origin, carrier) %>% 
  summarize( prop = sum(is.na(dep_delay)) / length(dep_delay)) %>% # summarize(mean(is.na(dep_delay))
  arrange(origin,desc(prop),carrier)
```

  


## Question 3

To investigate if there is any pattern of flight cancellations depending on the time of year, I first create a `canceled_by_doy` dataset with the proportion of canceled flights for each day of the year (`doy`) and `origin`:


```{r q3}
install.packages("lubridate")
# WIP
canceled_by_doy <- flights %>% 
  group_by(origin,year, month, day) %>% 
  summarise (prop = sum(is.na(dep_delay)) / length(dep_delay)) 
```


## Question 4

```{r}
# `q4` here
library(lubridate)
canceled_by_doy <- flights %>% 
    mutate(date = make_date(year,month,day)) %>% ##
  group_by(origin,year, month, day, date) %>% 
  summarise (prop = sum(is.na(dep_delay)) / length(dep_delay)) 

canceled_by_doy <- as.data.frame(canceled_by_doy)


ggplot(canceled_by_doy, aes(date, prop, colour = origin,apha=0.3)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm)
```
# Discussion
# From the table, you can see the EV carrier cancels flights significantly more often than the other carriers do when departing from the same origin in the tri-state area (JFK, LGA, EWR). The DL carrier cancels flights the least amount than the other carriers across all origin airports.

# In the plot, the shape of the curves (smoothened lines) are fairly linear and not steep for all origins throughout the year. Suggesting the average proportion of cancelled flights month to month departing from JFK, LGA, or EWR is fairly consistent . Though, it looks as though between February and March, there are a lot of flight cancellations departing from any of the three airports than there are at any other time in the year. 

# On the whole, it looks like throughout the year there are a lot of flight cancellations month to month for all origins but not as much as there are in the aforementioned time frames.
